Follow-up Samples

### 2016-12-07 submission/manuscript_outline_20161024.tex

@@

We then investigated which OTUs could potentially be more important in
our model {[}Figure 5 \& Table S4{]}. Many of the OTUs identified
classified to normal flora bacterium {[}Table S4{]}. Only a single OTU
though was significant after multiple comparison correction and the
lowest taxonomic identification of Otu000012 was to Blautia. Although
there was a difference in the relative abundance at initial and follow
up these values were not drastically different from the relative
abundance values observed in the control individuals of the training set
{[}Figure 5{]}. This research provides evidence that it is possible to
use bacterial microbiome data to create a highily sensitive model, that
is reactive to therapy, for detection of adenoma or carcinoma. It
accomplishes this by using a unique sample set in which before and after
surgery stool samples are available for assessment. By using these types
of samples we are not only able to show sensitivity of lesion prediction
but also able to show that this model is reactive. That is to say that
after surgery for removal of the adenoma or carcinoma it decreases the
positive probability to reflect a lower likelihood of the individual
having an adenoma or carcinoma.

@@

Naturally, it is curious that normal staples of many screening studies
such as Fusobacterium, Parvimonas, and Peptostreptococcus were not
present in the majority of the training models. One potential
explanation for this is that FIT provides the same information to the
model as these three organisms and so the model uses FIT preferentially
over them. This has been suggested to be the case in a previous study
(\textbf{insert Baxter Study}). Regardless, our study does not argue
against the importance of these bacterium in CRC initiation or
pathogenesis but rather that the model does not utilize these specific
bacteria for prediction purposes. Another potential reason why we did
not identify the ``usual suspects''" is that these bacteria may not
change much between initial and follow up samples in those with an
identified lesion. That is to say that the bacteria are consistently
present even after removal of the lesion by surgery. Finally, it is
likely that within our test set there was not enough indviduals in which
detection was made or relative abundance high enough for these bacteria
to be significant using a paired wilcoxson test.

@@

@@

>>>>

\textbf{\emph{Differences in Adenoma and Carcinoma in Previously
Associated Cancer Bacteria}} First, there was a clear magnitude
difference in these specific OTUs based on whether they were from
adenoma or carcinoma individuals {[}Figure 6{]}. The carcinoma samples
showed a significant difference between initial and follow up samples
for Peptostreptococcus assacharolytica (P-value = 0.0183) and
Porphyromonas stomatis (P-value = 0.0154) whereas there were no
signficant differences in any of these OTUs in the adenoma samples
{[}Table S5{]}.

@@

We then investigated which OTUs could potentially be more important in
our model {[}Figure 5 \& Table S4{]}. Many of the OTUs identified
classified to normal flora bacterium {[}Table S4{]}. Only a single OTU
though was significant after multiple comparison correction and the
lowest taxonomic identification of Otu000012 was to Blautia. Although
there was a difference in the relative abundance at initial and follow
up these values were not drastically different from the relative
abundance values observed in the control individuals of the training set
{[}Figure 5{]}. Although we were interested in what we could use to
classify those with either adenoma or carcinoma versus normal. We found
that the traditional bacteria associated with CRC were higher in
magnitude in the carcinoma group and there were significant differences
in some of these OTUs between the initial and follow up samples
{[}Figure 5 and Tabl S5{]}. This research provides evidence that it is
possible to use bacterial microbiome data to create a highily sensitive
model, that is reactive to therapy, for detection of adenoma or
carcinoma. It accomplishes this by using a unique sample set in which
before and after surgery stool samples are available for assessment. By
using these types of samples we are not only able to show sensitivity of
lesion prediction but also able to show that this model is reactive.
That is to say that after surgery for removal of the adenoma or
carcinoma it decreases the positive probability to reflect a lower
likelihood of the individual having an adenoma or carcinoma.

@@

Naturally, it is curious that normal staples of many screening studies
such as Fusobacterium, Parvimonas, and Peptostreptococcus were not
present in the majority of the training models. One potential
explanation for this is that FIT provides the same information to the
model as these three organisms and so the model uses FIT preferentially
over them. This has been suggested to be the case in a previous study
(\textbf{insert Baxter Study}). It is also possible that these specific
bacteria play a major role in the progression to carcinoma but may not
be as important in the initiation of an adenoma, which would be
supported by our data {[}Figure 5{]}. Regardless, our study does not
argue against the importance of these bacterium in CRC initiation or
pathogenesis but rather that the model does not utilize these specific
bacteria for prediction purposes. Another potential reason why we did
not identify the ``usual suspects''" is that these bacteria may not
change much between initial and follow up samples in those with an
identified lesion. That is to say that the bacteria are consistently
present even after removal of the lesion by surgery. Finally, it is
likely that within our test set there was not enough indviduals in which
detection was made or relative abundance high enough for these bacteria
to be significant using a paired wilcoxson test.

@@

\textbf{\emph{Investigation of Important CRC Bacteria}} In order to
investigate how specific bacteria that have been previously associated
with CRC with selected OTUs that taxonomically classified to
Fusobacterium Nucleatum, Parvimonas Micra, Peptostreptococcus
Assacharolytica, and Porphyromonas Stomatis. Specifically we wanted to
observe if there were any differences based on whether the individual
had an adenoma or carcinoma that may have been missed with our lesion
versus normal model.

@@

\textbf{Figure 6: Previously Associated CRC Bacteria in Initial and
Follow up Samples.} A) Carcinoma initial and follow up samples. There
was a significant difference in initial and follow up sample for the
OTUs classfied as Peptostreptococcus assacharolytica (P-value = 0.0183)
and Porphyromonas stomatis (P-value = 0.0154). B) Adenoma initial and
follow up samples. There were no significant differences between initial
and follow up.

### 2016-12-07 submission/supplemental_outline_20161024.Rmd

>>>>

**Table S5: Summary of P-values for Specific CRC OTUs between Initial and Follow Ups**

```{r supplemental_table5, echo=FALSE, message=FALSE, warning=FALSE}

pvalue_adn_crc_summary <- read.csv('../results/tables/adn_crc_maybe_pvalue_summary.csv', header = T, row.names = 1)

st5 <- cbind(tax_id = c("Fusobacterium Nucleatum", "Parvimonas Micra", 
                        "Peptostreptococcus Assacharolytica", "Porphyromonas Stomatis"), 
             crc_pvalue = format(pvalue_adn_crc_summary$crc_pvalue, digits = 3), 
             crc_bh = format(pvalue_adn_crc_summary$crc_BH, digits = 3), 
             adn_pvalue = format(pvalue_adn_crc_summary$adn_pvalue, digits = 3), 
             adn_bh = format(pvalue_adn_crc_summary$adn_BH, digits = 3))

colnames(st5) <- c("Taxonomic ID", "CRC P-value", "CRC BH Corrected P-value", 
                   "Adenoma P-value", "Adenoma BH Corrected P-value")

kable(st5, format = "markdown", align = 'c', padding = 2)
